Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-12 shanghai.tex
Reworked the section on restriction to the identity component.
Created a lemma to hold the proof that every discrete isogeny to G^0
extends to a discrete isogeny to G.
The proof that restriction to the identity component is surjective is
now very short, largely because it now uses the propositions regardling
finite geochars.

To define the category of finite geometrized characters we must work a bit harder.
Consider the following category \cdef{$C(G)$}:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ an isogeny (a finite, surjective etale morphism and a morphism of commutative group $\Fq$-schemes)
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G\\
\end{tikzcd}
\]
\Clifton{I suspect $\alpha$ is a covering, for free. Is that right? 
If $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, 
does it follow that $\alpha$ is also an etale cover? (Etale: Yes. Finite? Surjective?) 
If so, then I can simplify the definition of weak isomorphism accordingly.}
We say that $(\alpha,T)$ in $C(G)$ is a \cdef{weak isomorphism} if $\alpha$ is a covering and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

\David{This section needs to be modified once the bounded geometrized characters section is done}
Let $G^0$ be the connected component of $G$ containing the identity.
By Proposition~\ref{prop:pullback}, pull-back along
$G^0\hookrightarrow G$ determines a monoidal functor
$\cs{L}\mapsto \cs{L}\vert_{G^0}$ from $\GC(G)$ to $\GC(G^0)$.  In
this section we show that this functor is essentially surjective. To
do this we employ a perspective on characters sheaves which is
different from that in Section~\ref{sec:category} but which works
very well for geometrically connected abelian algebraic groups; this
perspective fails to extend properly to the more general context of
Section~\ref{sec:category}.

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component $G^0$ is a
  geometrically connected algebraic group over $\Fq$, clopen in $G$
  \cite{vdG&M}*{3.17}. By Proposition~\ref{prop:connected}, every
  geometrized character $\cs{E} = (\gcs{E},\mu^0,\phi^0)$ on $G^0$ descends to
  a line bundle on $G^0$.

  Let us review how line bundles on $G^0$ determine geometrized characters
  on $G^0$.  Up to isomorphism, line bundles on $G^0$ may be
  apprehended in terms of pairs $(\pi,\psi)$,
  \David{Aren't there line bundles which are not geometrized characters,
    for which $B$ is not an algebraic group?}
  where $\pi : B\to G^0$ is a finite etale morphism of algebraic
  groups and a character $\psi : A\to \EEx$, where
  $\Gal(\bFq/\Fq)$ acts trivially on $A \ceq \ker {\bar \pi}$, as
  summarized by the following diagram.
  \begin{equation}\label{E}
  \begin{tikzcd}
  \EEx & 1\rar & A \arrow[bend right,swap]{ll}{\psi} \rar & B \rar{\pi} & G^0 \rar & 1
  \end{tikzcd}
  \end{equation}
  (In fact, we may always take $B = G^0$ and $\pi = \Lang_{G^0}$, the
  Lang isogeny for $G^0$.)  Indeed, $\cs{E}$ is precisely the
  component of $\pi_! (\EE)_{B}$ on which $A$ acts through $\psi$.  To
  see that $\cs{E}$ is a line bundle on $G^0$, start by pulling back
  $\pi: B \to G^0$ along $m_0 : \GzxGz \to G^0$.
  \begin{equation}
  \begin{tikzcd}
  \EEx & 1\rar & A \arrow[bend right,swap]{ll}{\psi} \rar & B \rar{\pi} & G^0 \rar & 1\\
      & 1\rar & \AxA \arrow[dashed]{llu} \uar \rar
      & \BxB \rar{\pi\tight{1}{\times}{1} \pi} \uar{m_{B}} & \GzxGz \uar{m_0} \rar & 1
  \end{tikzcd}
  \end{equation}
  As pictured above, the kernel of this pull-back is $\AxA$;
  indeed, $\AxA \to \BxB$ is the pull-back of $A\to B$ along
  $m_{B}$. Use this Cartesian square to define a character (the dotted
  arrow, above) of $\AxA$; that is, define $\AxA \to \EEx$
  by the composition of $\AxA \to A$ and $\psi : A\to \EEx$.
  Note that this character is $\psi \otimes \psi : \AxA \to \EEx$.
  \Clifton{I haven't explained how maps between diagrams define sheaf
    homomorphisms between local systems.}
  Since the lower sequence is precisely the sequence defining
  $\ExE$, this shows that, by construction, $\cs{E}$
  comes equipped with a canonical isomorphism $m_0^*\cs{E} \iso \ExE$.
  Now the base change $\gcs{E}$ comes equipped with isomorphisms
  $\phi^0 : \Frob{G^0}^* \gcs{E} \to \gcs{E}$ and
  $\mu^0 : \bm_0^* \gcs{E} \to \bExE$
  satisfying conditions \ref{CS.1} and \ref{CS.2}. Then $(\gcs{E},\mu^0,\phi^0)$
  is a geometrized character of $G^0$.

  The next step is to show that $\pi: B \to G^0$ extends to a finite
  etale morphism of group schemes $f: H\to G$, smooth and locally of
  finite type over $\Fq$, in such a way that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G)
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.  Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to H^0(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces a map
  $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$,
  fitting into the long exact sequence derived from applying
  $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  $$\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).$$
  We now show that $\Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)) = 0$.

@@

  We therefore have the existence of diagram \eqref{extension-diagram}
  at the level of $\bFq$-points.  This expresses $H(\bFq)$ as a
  disjoint union of translates of $B(\bFq)$; by transport of structure
  we may take $H$ to be a group scheme over $\bFq$.  Similarly, the
  restriction of $f$ to each component of $H$ is a morphism of
  schemes, and thus $f$ is as well.  Finally, the whole diagram
  descends to a diagram of $\Fq$-schemes since the $\bFq$-points of
  the objects come equipped with continuous $\Gamma$-actions, and the
  morphisms are $\Gamma$-equivariant.

  Now we show that $f : H\to G$, as above, together with a character
  $\psi : \ker \bf \to \EEx$, defines a geometrized character on
  $G$. We will also show that the restriction of this geometrized character
  on $G$ to $G^0$ is precisely the geometrized character of $G^0$ defined by
  the pair $(f^0,\psi)$ as above.

  Set $X \ceq \pi_0(G)\times_{\Spec{\Fq}} \Spec{\bFq}$. We will
  identify $X$ with $\pi_0(H)\times_{\Spec{\Fq}} \Spec{\bFq}$ using
  the base change of $\pi_0(f): \pi_0(H)\to \pi_0(G)$. For each $x\in X$,
  write $\bf^x : \bH^x \to \bG^x$ for the
  restriction of $\bf$ to $\bH^x$, as a cover of
  $\bG^x$. For each $x,y\in X$, let
  $\bm_{x,y} : \bH^x \times \bH^{y} \to \bH^{x+y}$
  be the map obtained from $\bm$ by restriction. Define
  $\bm_{x} : A\to \Aut(\bf^x)$ by
  $\bm_{x}(a): g \mapsto \bm_{0,x}(a,g)$; then
  $\bm_{x} : A\to \Aut(\bf^x)$ is an isomorphism.
  \Clifton{I need to show that $\bm_{x} : A\to \Aut(\bf^x)$ is an isomorphism.}

  For each $x\in X$ let $\gcs{L}^x$ be the rank-$1$ local system on
  $\bG^x$ determined by the finite etale cover
  $\bf^x : \bH^x \to \bG^x$ and the character
  $\psi_x \ceq \psi \circ \bm_x^{-1}$.  Since
  $\bG = \coprod_{x\in X} \bG^x$ (disjoint union), this
  determines a constructible sheaf $\gcs{L}$ on $\bG$, locally
  constant of rank-$1$ on each connected component of $\bG$.
  Taking the case $x=0$ we observe that $\gcs{L}^0 =\gcs{E}$,
  the latter defined above. To finish the proof of the proposition we
  equip $\gcs{L}$ with isomorphisms
  $\mu : \bm^* \gcs{L} \to \LxL$ and
  $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ satisfying conditions \ref{CS.1}
  and \ref{CS.2}.

  Observe that $\bm: \bGxG \to \bG$ is
  completely determined by the family of maps
  $\bm_{x,y} : \bG^x \times \bG^{y} \to \bG^{x+y}$.
  Accordingly, to define $\mu$ for $\gcs{L}$ it is sufficient to define isomorphisms
  $\mu_{x,y} : \bm_{x,y}^*\gcs{L}^{x+y} \iso \gcs{L}^{x}\boxtimes \gcs{L}^{y}$
  for all $x,y\in X$. To do that, note first that the local system
  $\gcs{L}^{x}\boxtimes \gcs{L}^{y}$ is defined by the following diagram.
  \[
  \begin{tikzpicture}[auto,style={rectangle}]
  \node (0) at (0,0) {$\EEx$};
  \node (1) [anchor=east] at (6,0) {$\Aut(\bf^x) \times \Aut(\bf^y)$};
  \node (2) [anchor=west] at (6.4,0) {$\bH^x \times \bH^y$};
  \node (3) at (11,0) {$\bG^x \times \bG^y$};
  \draw [->] (1) to node[swap, label] {$\psi_x \otimes \psi_y$} (0);
  \draw [->] (6.3,-0.06) arc (330:40:0.12) -- (6.3,0.07);
  \draw [->] (2) to node[label] {$\bf^x \tight{0}{\times}{0} \bf^y$} (3);
  \end{tikzpicture}
  \]
  On the other hand, the diagram for the local system $\bm_{x,y}^* \gcs{L}^{x+y}$
  is found by pulling-back the diagram for $\gcs{L}^{x+y}$ along
  $\bm_{x,y} : \bH^x\times\bH^y \to \bH^{x+y}$;
  the result is the middle horizontal diagram below, where we use the
  fact that the covering group of the middle arrow is $\Aut(\bf^{x+y})$, canonically.

  \vspace{0.2in}

  \[
  \begin{tikzpicture}[auto,style={shape={asymmetrical rectangle}},scale=1.2]
  \node (12) [anchor=east] at (4,2.4) {$\Aut(\bf^{x+y})$};
  \node (22) [anchor=west] at (4.4,2.4) {$\bH^{x+y}$};
  \node (32) [anchor=east] at (11,2.4) {$\bG^{x+y}$};
  \node (01) at (0,1.2) {$\EEx$};
  \node (11) [anchor=east] at (4,1.2) {$\Aut(\bf^{x+y})$};
  \node (21) [anchor=west] at (4.4,1.2) {$\bH^{x+y} \times_{\bG^{x+y}} (\bG^x\times\bG^y)$};
  \node (31) [anchor=east] at (11,1.2) {$\bG^x\times \bG^y$};
  \node (10) [anchor=east] at (4,0) {$\Aut(\bf^x)\times \Aut(\bf^y)$};
  \node (20) [anchor=west] at (4.4,0) {$\bH^x\times \bH^y$};
  \node (30) [anchor=east] at (11,0) {$\bG^x\times \bG^y$};
  \draw [->] (4.3,2.34) arc (330:40:0.12) -- (4.3,2.47);
  \draw [->, densely dotted] (4.3,1.14) arc (330:40:0.12) -- (4.3,1.27);
  \draw [->] (4.3,-0.06) arc (330:40:0.12) -- (4.3,0.07);
  \draw [->] (12) to node[swap,label] {$\psi_{x+y}$} (01);
  \draw [->, dashed] (11) -- (01);
  \draw [->] (10) to node[label] {$\psi_x \otimes \psi_y$} (01);
  \draw [->, dashed] ($(11.south) + (0,-0.7)$) -- (11);
  \draw [double equal sign distance] (11) -- (12);
  \draw [->] (22) to node[label] {$\bf^{x+y}$} (32);
  \draw [->] ($(22.south) + (0,-0.7)$) -- (22);
  \draw [->] (21) -- (31);
  \draw [->] ($(32.225) + (0,-0.7)$) to node[swap] {\footnotesize{$\bm_{x,y}$}} (32.225);
  \draw [->] ($(22.south) + (0,-1.9)$) -- ($(22.south) + (0,-1.2)$);
  \draw [->] (20) to node[label] {$\bf^x \times \bf^y$} (30);
  \draw [double equal sign distance] ($(32.225) + (0,-1.9)$) -- ($(32.225) + (0,-1.2)$);
  \end{tikzpicture}
  \]
  Now, the homomorphism from the covering group of
  $\bf^x\times\bf^y$ to the covering group of
  $\bf^{x+y}$ induced by the composition
  $ \bH^x\times \bH^y \to \bH^{x+y} \times_{\bG^x\times\bG^y} \bG^x\times\bG^y \to \bH^{x+y}$
  is precisely $(\bm_x(a),\bm_y(b)) \mapsto \bm_{x+y}(ab)$.
  Since this map commutes with the characters $\psi_x\otimes\psi_y$
  and $\psi_{x+y}$ (so, the far-left triangle commutes in the diagram above)
  it follows that, for each $x,y\in X$ the diagram above defines a canonical isomorphism
  $\mu_{x,y} : \bm_{x,y}^* \gcs{L}^{x+y} \to \gcs{L}^x \boxtimes\gcs{L}^y$.
  This defines $\mu: \bm^* \gcs{L} \iso \LxL$.
  It is straightforward (but laborious) to show that the pair
  $(\gcs{L},\mu)$ satisfies condition \ref{CS.1}.

  Now we turn to the definition of $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$.
  As above, we define this isomorphism through a family of
  isomorphisms on the connected components of $\bG$. The
  restriction of $\Frob{G}$ to $\bG^x$ defines an isomorphism
  $\Frob{G}^x : \bG^x \to \bG^{\Frob{\Fq}(x)}$.
  %\[
  %\xymatrix{
  %\ar[d]_{\Frob_x} \bG^x \ar[r] & \ar[d]_{\Frob{G}} \bG\\
  %\bG^{\Frob{\Fq}(x)} \ar[r] & \bG\\
  %}
  %\]
  Pull back the finite etale covering in the diagram for
  $\gcs{L}^{\Frob{\Fq}(x)}$ along $\Frob{G}^x$ and compare the result with the
  finite etale covering in the diagram for $\gcs{L}^{x}$.

  \[
  \begin{tikzpicture}[auto,style={rectangle},scale=1.2]
  \node (12) [anchor=east] at (4,2.4) {$\Aut(\bf^{\Frob{\Fq}(x)})$};
  \node (22) [anchor=west] at (4.4,2.4) {$\bH^{\Frob{\Fq}(x)}$};
  \node (32) [anchor=west] at (8,2.4) {$\bG^{\Frob{\Fq}(x)}$};
  \node (01) at (0,1.2) {$\EEx$};
  \node (11) [anchor=east] at (4,1.2) {$\Aut(\bf^{\Frob{\Fq}(x)})$};
  \node (21) [anchor=west] at (4.4,1.2) {$\bH^{\Frob{\Fq}(x)} \times_{\bG^{\Frob{\Fq}(x)}} \bG^x$};
  \node (31) [anchor=west] at (8,1.2) {$\bG^x$};
  \node (10) [anchor=east] at (4,0) {$\Aut(\bf^x)$};
  \node (20) [anchor=west] at (4.4,0) {$\bH^x$};
  \node (30) [anchor=west] at (8,0) {$\bG^x$};
  \draw [->] (4.3,2.34) arc (330:40:0.12) -- (4.3,2.47);
  \draw [->, densely dotted] (4.3,1.14) arc (330:40:0.12) -- (4.3,1.27);
  \draw [->] (4.3,-0.06) arc (330:40:0.12) -- (4.3,0.07);
  \draw [->] (12) to node[swap,label] {$\psi_{\Frob{\Fq}(x)}$} (01);
  \draw [->, dashed] (11) -- (01);
  \draw [->] (10) to node[label] {$\psi_x$} (01);
  \draw [->, dashed] ($(11.south) + (0,-0.7)$) -- (11);
  \draw [double equal sign distance] (11) -- (12);
  \draw [->] (22) to node[label] {$\bf^{\Frob{\Fq}(x)}$} (32);
  \draw [->] ($(22.225) + (0,-0.7)$) -- (22.225);
  \draw [->] (21) -- (31);
  \draw [->] ($(32.222) + (0,-0.7)$) to node[swap] {\footnotesize{$\Frob{G}^x$}} (32.222);
  \draw [->] ($(22.225) + (0,-1.9)$) -- ($(22.225) + (0,-1.2)$);
  \draw [->] (20) -- (30);
  \draw [double equal sign distance] ($(32.222) + (0,-1.9)$) -- ($(32.222) + (0,-1.2)$);
  \end{tikzpicture}
  \]
  The covering group of $\bH^{\Frob{\Fq}(x)} \times_{\bG^{\Frob{\Fq}(x)}} \bG^x$
  is isomorphic to $\Aut(\bf^{\Frob{\Fq}(x)})$, canonically, so
  $(\Frob{G}^x)^* \gcs{L}^{\Frob{\Fq}(x)}$ is defined by the middle (horizontal)
  diagram, above.  We define the sheaf homomorphism
  $\phi_x : (\Frob{G}^x)^* \gcs{L}^{\Frob{\Fq}(x)} \to \gcs{L}^x$ by the canonical morphism
  $\bH^x \to \bH^{\Frob{\Fq}(x)} \times_{\bG^{\Frob{\Fq}(x)}} \bG^x \to $;
  to see that this defines a sheaf homomorphism it is sufficient to
  check that the character $\psi_x$ commutes with the family of isomorphisms
  $\phi_x : (\Frob{G}^* \gcs{L})\vert_{\bG^x} \iso (\Frob{G}^x)^* \gcs{L}^{\Frob{x}} \to \gcs{L}^x$
  now defines the isomorphism $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$.
  \Clifton{Add some detail here.}
  This isomorphism is compatible with $\mu$ in the sense of condition~\ref{CS.2}.
\end{proof}
\Clifton{I can flesh out this remark once I've promoted the definition of $\GCb(G)$ (category of bounded geometrized characters of $G$).}
\begin{remark}
  We have seen that if $f : H\to G$ is a finite etale cover of $G$ by
  a smooth group scheme and if all geometric automorphisms of $f$ are
  fixed by Frobenius, then every character of $\ker f$ determines a
  geometrized character of $G$. However, it is not true that every character
  sheaf on $G$ arises in this manner.
\end{remark}

%\input{centralextensions}

>>>>

To define the category of finite geometrized characters we must work a bit harder.
Consider the following category \cdef{$C(G)$}:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is a \cdef{discrete isogeny} (a finite, surjective etale morphism and a morphism of commutative group $\Fq$-schemes, for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $f^{-1}{\bar e}$ is trivial), 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G\\
\end{tikzcd}
\]
\Clifton{I suspect $\alpha$ is a covering, for free. Is that right? 
If $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, 
does it follow that $\alpha$ is also an etale cover? (Etale: Yes. Finite? Surjective?) 
If so, then I can simplify the definition of weak isomorphism accordingly.}
We say that $(\alpha,T)$ in $C(G)$ is a \cdef{weak isomorphism} if $\alpha$ is a covering and $T$ is an isomorphism of vector spaces.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.

@@

Let $G^0$ be the connected component of $G$ containing the identity.
By Proposition~\ref{prop:pullback}, pull-back along
$G^0\hookrightarrow G$ determines a monoidal functor
$\cs{L}\mapsto \cs{L}\vert_{G^0}$ from $\GC(G)$ to $\GC(G^0)$.  In
this section we show that this functor is essentially surjective.

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component, 
  $G^0$, is a connected algebraic group over $\Fq$, 
  clopen in $G$ \cite{vdG&M}*{3.17}. 
  By Proposition~\ref{prop:connected}, every
  geometrized character on $G^0$ is isomorphic to a 
  finite geometrized character on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite geochar on $G^0$ extends to a geochar on $G$.

 Let $(\pi,\psi)$ be a finite geochar on $G^0$.
 By Lemma~\ref{lemma:ext}, there is an extension of the 
 descrete isogeny $\pi : B \to G^0$ 
 (see Section~\ref{sec:finite} for the definition of a discrete isogeny) 
 to a discrete isogeny $f : H \to G^0$ 
 such that $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism.
 Then $(f,\psi)$ is a finite geochar on $G$.
 Then $(f,\psi)\vert_{G^0} \iso (\pi,\psi)$,
 completing the proof that $\GC(G)\to \GC(G^0)$ is essentially surjective.
\end{proof}

\begin{lemma}\label{lemma:ext}
Every discrete isogeny onto $G^0$ extends to a discrete isogeny onto $G$ such that the component group of the covering space is isomorphic to the component group of $G$.
\end{lemma}

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A = \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G)
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

  Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to H^0(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$,
  fitting into the long exact sequence derived from applying
  $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  $$\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).$$
  We now show that $\Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)) = 0$.

@@

  We therefore have the existence of diagram \eqref{extension-diagram}
  at the level of $\bFq$-points.  This expresses $H(\bFq)$ as a
  disjoint union of translates of $B(\bFq)$; by transport of structure
  we may take $H$ to be a group scheme over $\bFq$.  Similarly, the
  restriction of $f$ to each component of $H$ is a morphism of
  schemes, and thus $f$ is as well.  Finally, the whole diagram
  descends to a diagram of $\Fq$-schemes since the $\bFq$-points of
  the objects come equipped with continuous $\Gamma$-actions, and the
  morphisms are $\Gamma$-equivariant.
\end{proof}
